### 2022-03-13 papers/sbst22/paper_sbst22.tex
Add some text based on logs

\documentclass[sigconf,review,anonymous]{acmart}

\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{listings}
\usepackage{algorithmic}
\usepackage{float}
%\usepackage[numbers,sort&compress,square]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{subcaption}
%\usepackage{hyperref}
\usepackage{color}
%\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{soul}
\usepackage{xspace}
\usepackage{boxedminipage}
\usepackage{alltt}
\usepackage{multirow}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{balance}
\definecolor{light-gray}{gray}{0.90}

\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}
\floatname{algorithm}{Algorithm}

\newtheorem{definition}{Definition}

 %krams hinter fontadjust ist neu
  \definecolor{lightgrey}{rgb}{0.90,0.90,0.90}
\lstset{escapeinside={(*}{*)}}
  \lstloadlanguages{java}
 \lstdefinelanguage{pseudocode}
  {morekeywords={if, else, initialize, return, for, each, in, global, new}
   }
  \lstset{
    tabsize=2,
    mathescape=true,
    escapeinside={(*}{*)},
    captionpos=t,
    framerule=0pt,
    backgroundcolor=\color{lightgrey},
    basicstyle=\scriptsize\ttfamily,
    keywordstyle=\footnotesize\bfseries,
    numbers=none,
    numberstyle=\tiny,
    numbersep=1pt,
    fontadjust,
    breaklines=true,
    breakatwhitespace=false
  }

% \hypersetup{
% colorlinks=true,
% urlcolor=rltblue,
% linkcolor=rltred,
% citecolor=rltgreen,
% bookmarksnumbered=true,
% pdftitle={EvoSuite at the SBST 2016 Tool Competition},
% pdfauthor={Gordon Fraser and Andrea Arcuri},
% pdfsubject={Test case generation},
% pdfkeywords={Test case generation, unit testing, test
%   oracles, assertions, search based testing}
% }

\definecolor{rltred}{rgb}{0.5,0,0}
\definecolor{rltgreen}{rgb}{0,0.5,0}
\definecolor{rltblue}{rgb}{0,0,0.5}
\definecolor{ScarletRed}{rgb}{0.80,0.00,0.00}

% in draft mode we put \remarks into the margins and do other stuff
% set to \draftfalse for 
\newif\ifdraft
\draftfalse

\ifdraft
	\marginparwidth=1.3cm
	\marginparsep=5pt
	\newcommand\remark[1]{%
		\mymarginpar{\raggedright\hbadness=10000\tiny\it #1\par}}
	% TODO marker
	\newcommand{\TODO}[1]{\sethlcolor{yellow}\textbf{\textcolor{ScarletRed}{\hl{TODO: #1}}}\xspace}
\else
	\newcommand\remark[1]	{}
	\newcommand{\TODO}[1]{}
\fi

\ifdraft
	\overfullrule3pt
\fi

% We use \FIXME for located problems (``defect'')
\newcommand{\FIXME}[1]{\remark{FIXME: #1}}
\newcommand\parremark[1]	{\par\textbf{REMARK:} #1\par}

\newcommand{\gordon}[1]{\textcolor{blue}{\sf\small\textbf{Gordon:} #1}}
\newcommand{\andrea}[1]{\textcolor{ScarletRed}{\sf\small\textbf{Andrea:} #1}}

% \mathid is used to denote identifiers and slots in formulas
\newcommand{\mathid}[1]{\text{\rmfamily\textit{#1}}}

% But usually, we shall use \|name| instead.
\def\|#1|{\mathid{#1}}

% \codeid is used to denote computer code identifiers
\newcommand{\codeid}[1]{\texttt{#1}}

% But usually, we shall use \<name> instead.
\def\<#1>{\codeid{#1}}

% Our results
\newenvironment{result}%
{\smallskip
\noindent
\let\emph=\textbf
\begin{boxedminipage}{\columnwidth}\begin{center}\em}%
{\end{center}\end{boxedminipage}%
\smallskip
}

\newcommand{\JodaTime}{Joda-Time\xspace}  % That's how they write themselves -- AZ

\newcommand{\EVOSUITE}{\textsc{EvoSuite}\xspace}
\newcommand{\JTEXPERT}{\textsc{jTExpert}\xspace}
\newcommand{\RANDOOP}{\textsc{Randoop}\xspace}
\newcommand{\TT}{\textsc{T3}\xspace}

\newcommand{\MUTEST}{{\sc $\mu$Test}\xspace}
\newcommand{\CS}{{\sc SF100}\xspace}

\DeclareMathSymbol{,}{\mathpunct}{letters}{"3B}
\DeclareMathSymbol{,}{\mathord}{letters}{"3B}
\DeclareMathSymbol{\decimal}{\mathord}{letters}{"3A}
%%%"

\usepackage{siunitx}
\input{data/macros}

%------------------------------------------------------------------------- 
% take the % away on next line to produce the final camera-ready version 
%\pagestyle{empty}

%------------------------------------------------------------------------- 
\begin{document}

%

%\title{Unit Testing Tool competition: Results for EvoSuite}
\title{\EVOSUITE at the SBST 2022 Tool Competition}

\author{%
%
\IEEEauthorblockN{%
Sebastian Schweikl\IEEEauthorrefmark{1},
Gordon Fraser\IEEEauthorrefmark{1}, 
Andrea Arcuri\IEEEauthorrefmark{2},
%Jose Campos\IEEEauthorrefmark{3},
%and Annibale Panichella\IEEEauthorrefmark{4}%
}
%
\IEEEauthorblockA{\IEEEauthorrefmark{1}University of Passau, Germany\\
\{sebastian.schweikl,gordon.fraser\}@uni-passau.de}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Kristiania University College and Oslo Metropolitan University, Norway\\andrea.arcuri@kristiania.no}
\IEEEauthorblockA{\IEEEauthorrefmark{3}University of Lisbon, Portugal\\jcmcampos@fc.ul.pt}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Delft University of Technology, Netherlands\\A.Panichella@tudelft.nl}%
}

%\thispagestyle{empty}

\begin{abstract}
  \EVOSUITE is a search-based unit test generation tool for Java. This paper summarises the results and experiences of \EVOSUITE's participation at the 10th unit testing competition at SBST 2022, where \EVOSUITE achieved the highest overall score.
\end{abstract}

\maketitle

%------------------------------------------------------------------------- 
\section{Introduction}
%
For the last 10 years, the annual unit test generation competition has
evaluated progress on unit test generation tools using a benchmark of Java
classes, on which different tools are applied and compared. In the 2022 edition
of the competition, seven tools were applied on a set of \cuts classes. Details
about the procedure of the competition, the technical framework, and the
benchmark classes can be found in the competition
report~\cite{SBST-toolcomp22}. \EVOSUITE achieved an overall score of \score,
which was the highest among the competing and baseline tools.

%------------------------------------------------------------------------- 
\section{About EvoSuite}

\EVOSUITE~\cite{FrA11c} automatically generates test suites for Java classes
using evolutionary search. As input \EVOSUITE only requires the name of the
target class as well as the classpath on which the compiled bytecode of the
target class as well as its dependencies are located. \EVOSUITE first extracts
information about the relevant classes as well as their constructors, methods,
and fields using a basic static analysis. It instruments the bytecode while
classes are loaded by the Java classloader, such that \EVOSUITE can produce
execution traces for test executions, and to avoid test flakiness by replacing
non-deterministic calls with deterministic, mocked versions. Finally, \EVOSUITE
applies meta-heuristic search algorithms to automatically produce JUnit test
cases aimed at maximising code coverage.
%Besides the basic command line interface there are also plugins for popular development environments such as IntelliJ, Eclipse, or Maven~\cite{ICST16_Tool}.

\EVOSUITE generates tests using evolutionary algorithms. 
%
For this, it uses a genetic encoding of test cases consisting of variable-length sequences of Java statements (e.g., primitive statements as well as calls on constructors or methods). Standard evolutionary search operators (e.g., selection, crossover, mutation) are adapted for this representation. 
%
\EVOSUITE provides several different evolutionary algorithms to operate on this
representation. The initial approach was to optimise entire test suites with
respect to their overall code coverage~\cite{GoA_TSE12}. However, after
extensive evaluation~\cite{ea_evaluation} the default is now the
\textit{Dynamic Many-Objective Sorting Algorithm} (DynaMOSA) search
algorithm~\cite{dynamosa}, which operates at the test case level.
%
The search is driven by fitness functions that encode coverage criteria. \EVOSUITE can optimise for several criteria simultaneously, and the core fitness functions in \EVOSUITE are based on traditional heuristics for code coverage, such as the branch distance and the approach level (see~\cite{GoA_TSE12} for more details). However, there are several further fitness functions based on mutation testing~\cite{emse14_mutation} as well as other basic criteria~\cite{rojas2015combining}.

A large share of the complexity of \EVOSUITE's code base is due to efforts to
avoid flakiness in the generated tests. One the one hand, this covers extensive
instrumentation and mocking. On the other hand, \EVOSUITE also compiles and
executes each test individually to avoid compilation errors (which may be the
result of bugs in \EVOSUITE) or flakiness caused by non-determinism in the
class under test.

\EVOSUITE is also heavily optimised in order to produce tests that are as readable as possible and avoid test smells~\cite{panichella2020revisiting}. For this, \EVOSUITE applies post-processing once the available search budget has been exhausted~\cite{FrA11c,FrA13a}, consisting of minimisation to remove redundant tests and statements, and adding a minimised set of assertions  using mutation analysis~\cite{10.1109/TSE.2011.93}. 
%Besides multiple in-depth empirical evaluations of \EVOSUITE with respect to
%code coverage~\cite{fraser2014large,emse_archive,ea_evaluation,dynamosa},
%fault-finding effectiveness~\cite{shamshiri2015automatically,moein2017}, and
%effects on developer productivity~\cite{TOSEM_userstudy,ISSTA15_Study} and
%software maintenance~\cite{ICST2018_Maintenance}, \EVOSUITE has also participated in all prior instances of the unit testing tool competition, usually ranking among the best performing tools.

%------------------------------------------------------------------------- 
\section{Competition Setup}

The 2022 competition setup was identical to the 2021 setup: The most recent
version of \EVOSUITE (1.1.0) was used, which by default uses
DynaMOSA~\cite{dynamosa} as the search algorithm. No new features were added in
this version, but support for Java 9+ was added, several bugs were fixed, and a
major refactoring was applied to improve the code quality with respect to the
use of Java generics internally. While several smaller bugfixes were applied in
the run-up to the 2021 competition, no further optimisation or testing was
applied for 2022.

The configuration is largely identical to the established configuration used in
prior years, and mainly uses the tuned default
values~\cite{arcuri2013parameter} and default coverage criteria
\cite{rojas2015combining}.
%
As in previous years, the configuration omits the test minimisation step as the
competition score does not reward test minimality, and the minimisation takes
substantial computational effort. Similarly, the configuration includes all
regression assertions rather than filtering them with mutation analysis to save
time.
%
The time budget was again distributed by allocating 50\% of the overall time
set by the competition organisers for the search, and the other 50\% equally to
the remaining phases (e.g., assertion generation, compile-check,
flakiness-check).

%------------------------------------------------------------------------- 
\section{Results}

%\begin{figure}
%  \centering
%  \includegraphics[width=\columnwidth]{data/CoverageBoxV}
%  \caption{Coverage results achieved in the competition.}
%  \label{fig:results}
%\end{figure}

\begin{figure*}
  \centering
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\linewidth]{data/CoverageByBenchmark30.pdf}
    \caption{Coverage results achieved in the competition with \SI{30}{\second}.}
    \label{fig:results30}
  \end{subfigure}
  \hfill
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\linewidth]{data/CoverageByBenchmark120.pdf}
    \caption{Coverage results achieved in the competition with \SI{120}{\second}.}
    \label{fig:results120}
  \end{subfigure}

  \caption{Creating subfigures in \LaTeX.}
  \label{fig:figures}
\end{figure*}

With an overall score of \score, \EVOSUITE achieved the highest score of all
tools in the competition. As described in the competition
report~\cite{SBST-toolcomp22}, the score is calculated on the subset of \cuts classes
for which all tools produced tests in some of the runs.
Figure~\ref{fig:results30} summarises the results of \EVOSUITE in terms of the
mean coverage for the \cuts classes used in the competition for the \budgetShort search budget; Figure~\ref{fig:results120} shows the results for the \budgetLong search budget. While the coverage
increased from the \budgetShort to the \budgetLong search budget (mean line
coverage of \avgLinesCoverageRatioShort vs. \avgLinesCoverageRatioLong, respectively),
the additional run of 300 seconds
actually shows a slight decrease (mean line coverage of 73.9\%). 
%
The likely reason for this result is that \EVOSUITE ran out of time more
frequently when given more time. Indeed unlike past instances of the
competition we found no crashes of \EVOSUITE (e.g., exceptions thrown by
\EVOSUITE) at all, but in all cases where no tests were generated the \EVOSUITE
process had been killed by the competition setup, suggesting that \EVOSUITE was
using more time than allocated.

While the results for line coverage and branch coverage are similar, the
mutation scores are substantially lower. The mutation scores on the Guava
classes are highest compared to the other three projects from which classes
were taken; this is surprising since Guava has been known for years to be
notoriously challenging for
\EVOSUITE~\cite{campos2015continuous,almulla2017using}. However, we find that
in many cases the mutation analysis failed because dependency classes were
missing on the classpath. While \EVOSUITE, JUnit and the coverage measurement
used in the competition apparently were able to cope with this problem anyway,
the PIT mutation analysis tool was unable to deal with this problem. This may
likely be an error in the competition setup, but it would affect all test
generation tools. While Spoon classes are affected the most, similar problems
can be observed for all projects: Considering the \budgetLong runs we found
\texttt{ClassNotFoundException}s in 414 runs for Spoon, 309 for Guava, 308 for
Fastjson, and 329 for classes in Seata; in all these cases PIT was unable to
run even when it was possible to measure coverage.

%------------------------------------------------------------------------- 
\section{Conclusions}

This paper reports on the participation of the \EVOSUITE test generation tool
in the tenth SBST Java Unit Testing Tool Contest. On average, \EVOSUITE
achieved \avgConditionsCoverageRatioLong branch coverage,
\avgLinesCoverageRatioLong line coverage, and a mutation score of
\avgMutantsCoverageRatioLong, using a search budget of \budgetLong on the \cuts
classes considered for the competition. Overall, this results in a score of
\score, which is the highest score of all tools in the competition.

To learn more about \EVOSUITE, visit our Web site:
\begin{center}
%\url{http://evosuite.org/}
\texttt{http://www.evosuite.org}
\end{center}

%-------------------------------------------------------------------------

%\noindent
\textbf{Acknowledgments:} Many thanks to all the contributors to
\EVOSUITE.
%This project has been supported by EPSRC project % ``GREATEST'' 
%EP/N023978/2.

%------------------------------------------------------------------------- 
%\def\IEEEbibitemsep{5pt plus 1pt}
%\def\IEEEbibitemsep{6pt}
%\clearpage
\bibliographystyle{IEEEtranS}
\bibliography{papers}
\balance

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: